### 2023-06-05 text/paper.tex
overleaf

% Added link to preamble
\input{preamble}
\addbibresource{}

\title{Names diffusion and social class}
\author{}
\date{May 2023}

\begin{document}

\maketitle

\section{Introduction}

\begin{figure}[H]
    \centering
    \includegraphics[width=13cm]{plot/new_names_year.png}
    \caption{Number of new names per year in Chile (1920-2022)}
    \label{}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=13cm]{plot/name_commune.png}
    \caption{top names per commune in Chile (1920-2022)}
    \label{}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=13cm]{plot/names_commune_network.png}
    \caption{top names per commune in Chile (1920-2022)}
    \label{}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=13cm]{plot/top_nombre_comuna1.png}
    \caption{top names per commune in Chile (1920-2022)}
    \label{}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=13cm]{plot/top_nombre_comuna2.png}
    \caption{top names per commune in Chile (1920-2022)}
    \label{}
\end{figure}

\section{Methodology}

\end{document}